\begin{filecontents*}{references.bib}
	@article{lelli2016sparc,
		author  = {Lelli, Federico and McGaugh, Stacy S. and Schombert, James M.},
		title   = {SPARC: Mass Models for 175 Disk Galaxies with Spitzer Photometry at 3.6 $\mu$m},
		journal = {The Astronomical Journal},
		year    = {2016},
		volume  = {152},
		number  = {6},
		pages   = {157},
		doi     = {10.3847/0004-6256/152/6/157}
	}
	@article{mcgaugh2016rar,
		author  = {McGaugh, Stacy S. and Lelli, Federico and Schombert, James M.},
		title   = {Radial Acceleration Relation in Rotationally Supported Galaxies},
		journal = {Physical Review Letters},
		year    = {2016},
		volume  = {117},
		pages   = {201101},
		doi     = {10.1103/PhysRevLett.117.201101}
	}
	@article{meidt2014mlr36,
		author  = {Meidt, S. E. and Schinnerer, E. and van de Ven, G. and others},
		title   = {Reconstructing the Stellar Mass Distributions of Galaxies from S$^4$G IRAC 3.6 $\mu$m Images: Implications for the Stellar Mass-to-Light Ratio},
		journal = {The Astrophysical Journal},
		year    = {2014},
		volume  = {788},
		number  = {2},
		pages   = {144},
		doi     = {10.1088/0004-637X/788/2/144}
	}
	@article{schombert2014cmlr,
		author  = {Schombert, James M. and McGaugh, Stacy S.},
		title   = {Color--Mass-to-Light Ratio Relations for Disk Galaxies},
		journal = {The Astronomical Journal},
		year    = {2014},
		volume  = {148},
		number  = {4},
		pages   = {77},
		doi     = {10.1088/0004-6256/148/4/77}
	}
	@article{dutton2014mcr,
		author  = {Dutton, Aaron A. and Macci{\`o}, Andrea V.},
		title   = {Cold Dark Matter Haloes in the Planck Era: Concentration--Mass Relation},
		journal = {Monthly Notices of the Royal Astronomical Society},
		year    = {2014},
		volume  = {441},
		number  = {4},
		pages   = {3359--3374},
		doi     = {10.1093/mnras/stu742}
	}
	@article{planck2018vi,
		author  = {{Planck Collaboration} and Aghanim, N. and Akrami, Y. and others},
		title   = {Planck 2018 Results. VI. Cosmological Parameters},
		journal = {Astronomy \& Astrophysics},
		year    = {2020},
		volume  = {641},
		pages   = {A6},
		doi     = {10.1051/0004-6361/201833910}
	}
	@article{rahvar2014_nonlocal_rotation,
		author  = {Rahvar, S. and Mashhoon, B.},
		title   = {Observational Tests of Nonlocal Gravity: Galaxy Rotation Curves and Clusters of Galaxies},
		journal = {Physical Review D},
		year    = {2014},
		volume  = {89},
		pages   = {104011},
		doi     = {10.1103/PhysRevD.89.104011}
	}
	@article{cisneros2014_lcm,
		author  = {Cisneros, S. and Oblath, N. S. and Formaggio, J. A. and others},
		title   = {The Luminous Convolution Model as an Alternative to Dark Matter in Spiral Galaxies},
		journal = {arXiv e-prints},
		year    = {2014},
		eprint  = {1407.7583},
		archivePrefix = {arXiv}
	}
	@article{milgrom1983mond,
		author = {Milgrom, Mordehai},
		title = {A modification of the Newtonian dynamics as a possible alternative to the hidden mass hypothesis},
		journal = {Astrophysical Journal},
		year = {1983},
		volume = {270},
		pages = {365--370},
		doi = {10.1086/161130}
	}
	@article{bekenstein2004teves,
		author = {Bekenstein, Jacob D.},
		title = {Relativistic gravitation theory for the MOND paradigm},
		journal = {Physical Review D},
		year = {2004},
		volume = {70},
		pages = {083509},
		doi = {10.1103/PhysRevD.70.083509}
	}
	@article{famaey2012mondreview,
		author = {Famaey, Benoit and McGaugh, Stacy S.},
		title  = {Modified Newtonian Dynamics (MOND): Observational Phenomenology and Relativistic Extensions},
		journal = {Living Reviews in Relativity},
		year = {2012},
		volume = {15},
		pages  = {10},
		doi    = {10.12942/lrr-2012-10}
	}
	@article{FamaeyMcGaugh2012,
		author  = {Famaey, Benoit and McGaugh, Stacy S.},
		title   = {Modified Newtonian Dynamics (MOND): Observational Phenomenology and Relativistic Extensions},
		journal = {Living Reviews in Relativity},
		year    = {2012},
		volume  = {15},
		pages   = {10},
		doi     = {10.12942/lrr-2012-10}
	}
	@article{Bekenstein2004,
		author  = {Bekenstein, Jacob D.},
		title   = {Relativistic Gravitation Theory for the MOND Paradigm},
		journal = {Phys. Rev. D},
		year    = {2004},
		volume  = {70},
		pages   = {083509},
		doi     = {10.1103/PhysRevD.70.083509}
	}
	@software{H1_repo_2025,
  author    = {Figueiredo, Vítor M. F.},
  title     = {Baryon-Convolved Effective Potential (H1)},
  year      = {2025},
  publisher = {Zenodo},
  version   = {v0.1.0-rc},
  doi       = {10.5281/zenodo.16967259},
  url       = {https://doi.org/10.5281/zenodo.16967259}
}

\end{filecontents*}

\documentclass[11pt,a4paper]{article}

% ---------- Packages ----------
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{siunitx}
\emergencystretch=2em
\DeclareSIUnit\kpc{kpc}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{hyperref}
\usepackage[nameinlink,capitalise]{cleveref}


% ---------- Hyperref ----------
\hypersetup{
	colorlinks=true,
	linkcolor=blue!50!black,
	citecolor=teal!60!black,
	urlcolor=magenta!60!black,
	pdftitle={An Evolving Effective Potential Model for the Mass Discrepancy in Galaxies},
	pdfauthor={Vitor M. F. Figueiredo - vitor.figueiredo.research@protonmail.com}
}

% ---------- Simple units/macros ----------
\newcommand{\Msol}{M_{\odot}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\vect}[1]{\bm{#1}}
\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}

% Model symbols
\newcommand{\Lmem}{L}       % sample-global kernel range [kpc]
\newcommand{\gain}{\mu}     % global gain (dimensionless)
\newcommand{\rhob}{\rho_b}  % baryonic density
\newcommand{\phieff}{\Phi_{\mathrm{eff}}}
\newcommand{\geff}{\vect{g}_{\mathrm{eff}}}

% Boxed callouts
\newcommand{\keybox}[1]{%
	\vspace{0.6em}\noindent
	\colorbox{gray!10}{\parbox{\dimexpr\linewidth-2\fboxsep\relax}{#1}}%
	\vspace{0.6em}
}

% =========================================================
\begin{document}
	
	\begin{center}
		{\LARGE \textbf{A Baryon-Convolved Effective Potential for the Mass Discrepancy in Disk Galaxies}}\\[0.6em]
		{\large Vitor M. F. Figueiredo}\\
		\texttt{ORCID: 0009-0004-7358-4622 \quad (Vitor.figueiredo.research@protonmail.com)}\\[0.5em]
		\today
	\end{center}
		
		\begin{abstract}
			I introduce a \emph{sample-global two-parameter} baryon-built effective potential for disk galaxies. The total field is $\geff=-\nabla\phieff$, with
			\[
			\phieff(\vect{r})=\Phi_b(\vect{r})
			+\gain\,G\!\int \rhob(\vect{r}')\,U\!\big(\lVert\vect{r}-\vect{r}'\rVert;\Lmem\big)\,\dd^3\vect{r}'.
			\]
			Here $\Lmem$ (\si{\kpc}) and $\gain$ (dimensionless) are \emph{global across the sample}. “Evolving” indicates that $\phieff$ inherits a system’s formation history through its current baryon map; no explicit time dependence is fitted in H1. I benchmark rotation curves against a parity-matched halo baseline, verify BTFR/RAR behavior, and predict galaxy--galaxy lensing without refitting, with hard preregistered fail thresholds.
		\end{abstract}
		
		\section{Introduction}\label{sec:intro}
		The disk-galaxy mass discrepancy is usually addressed with per-galaxy dark halos. Here I test a parsimonious alternative: a cored, finite kernel convolved with the observed baryons to build an \emph{effective potential} with just two \emph{sample-global} parameters $(\Lmem,\gain)$, then ask if the same field explains RCs, BTFR, RAR, and out–of–sample lensing.
		\paragraph{Visual intuition (laser and smoke).}
		Think of the galaxy’s baryons as a three dimensional “smoke” (stars, dust, gas, circumgalactic gas). The bright thin disk is like a horizontal laser sheet slicing the smoke: star formation lights up near the mid--plane, while mass outside the activation layer remains dim. The convolution that defines $\Phi_{\rm eff}$ integrates the \emph{entire} smoke to build one static, sample-global field used for both rotation and lensing.
		\paragraph{Scope of “evolving” and core postulate.}
		Here “evolving” means that $\Phi_{\rm eff}$ \emph{inherits} a system’s formation history through the present--day baryon map $\rho_b$; no explicit time dependence is fitted in H1. Our core empirical postulate is \emph{dual--role consistency}: the same scalar potential $\Phi_{\rm eff}$ governs stellar kinematics via $-\nabla\Phi_{\rm eff}$ and sources lensing via the Newtonian Poisson closure $\nabla^2\Phi_{\rm eff}=4\pi G(\rho_b+\rho_{\rm eff})$. This closure is not derived from a fundamental theory here; it is a testable modeling hypothesis adjudicated by the preregistered gates.
		
		\section{Model: one potential for kinematics and lensing}\label{sec:model}
		\paragraph{Definition (scalar form).}
		
		\begin{align}
			\phieff(\vect{r}) &= \Phi_b(\vect{r})
			+\gain\,G\!\int \rhob(\vect{r}')\,U\!\big(\lVert\vect{r}-\vect{r}'\rVert;\Lmem\big)\,\dd^3\vect{r}', \label{eq:phi_eff}\\
			\geff(\vect{r}) &= -\nabla\phieff(\vect{r}). \label{eq:geff}
		\end{align}
		The radial derivative $K(r;\Lmem)\equiv U'(r;\Lmem)$ is a \emph{cored} acceleration kernel. I test two finite, positive choices:
		\begin{align}
			K_{\text{Plummer}}(r;\Lmem) &= \frac{1}{4\pi\,(r^2+\Lmem^2)}, \label{eq:Kpl}\\
			K_{\exp\text{-core}}(r;\Lmem) &= \frac{e^{-r/\Lmem}}{4\pi\,(r^2+\Lmem^2)}. \label{eq:KexpCore}
		\end{align}
		For \eqref{eq:Kpl}, $U(r)=(4\pi\Lmem)^{-1}\arctan(r/\Lmem)$ with $U(0)=0$. For $K_{\exp\text{-core}}$ I evaluate the scalar kernel via its 3D Fourier transform, i.e., I use $\widehat{U}(k)$ numerically; a closed–form $U(r)$ is not required for our FFT pipeline.
		
		\paragraph{Kinematics.}
		For an axisymmetric disk, $v^2(R)=R\,\lVert\geff(R,0)\rVert$ in the mid--plane.
		
		\paragraph{Tails and scaling.}
		I deliberately avoid $K\propto 1/r$ (which hard-wire exactly flat asymptotes) to keep BTFR scaling under control; quasi--flat outer segments are then an empirical outcome over finite radii and are adjudicated by our gates.
		
\keybox{\textbf{Single-field statement.}\;\;I compute one scalar potential $\Phi_{\rm eff}$ from baryons and use it for both kinematics and lensing. Stars trace $-\nabla\Phi_{\rm eff}$. For lensing I form $\rho_{\rm eff}=\nabla^2\Phi_{\rm eff}/(4\pi G)$ and project to $\Sigma$ and $\Delta\Sigma$ (App.~\ref{app:lensing}).}

		
		\paragraph{Lensing postulate.}
		I adopt as a modeling postulate that the same scalar $\Phi_{\rm eff}$ governing kinematics sources weak lensing via the Poisson mapping $\rho_{\rm eff}=\nabla^2\Phi_{\rm eff}/(4\pi G)$ and line–of–sight projection (App.~\ref{app:lensing}).
		\paragraph{Interpretation.} $\rho_{\rm eff}$ is an \emph{effective} source obtained by inverting Poisson’s equation for the modeled $\Phi_{\rm eff}$; local sign changes (compensation) can occur and are not interpreted as literal negative mass. Lensing is computed from $\Phi_{\rm eff}$ itself (App.~\ref{app:lensing}).
		
		\section{Numerical Implementation (simulation-ready)}\label{sec:numerics}
		\subsection{Grids and padding}\label{subsec:grids}
		Each galaxy is embedded in $(x,y,z)\in[-R_{\max},R_{\max}]^2\times[-Z_{\max},Z_{\max}]$ with \emph{padding factor 2} per dimension to suppress FFT wrap-around. The kernel is truncated at $R_{\rm trunc}=6\,\Lmem$ (tail contribution $<3\%$ beyond). Grid sizing scales with $\Lmem$: I enforce $R_{\max}\ge 1.5\,R_{\rm trunc}$ so the truncated kernel fits within the padded box for all fitted $L$. The baryon density $\rhob$ stacks stellar disk+bulge (SPARC $3.6\,\mu$m) and gas; scale heights follow catalog values or a thin-disk default.
		
		\subsection{FFT recipe: potential then gradient}\label{subsec:fft}
		Compute $\widehat{U}(\vect{k};\Lmem)$, then
		\[
		\widehat{\Phi}_K(\vect{k})=\gain\,G\,\widehat{\rhob}(\vect{k})\,\widehat{U}(\vect{k}),
		\qquad
		\widehat{\vect{g}}_K(\vect{k})=i\,\vect{k}\,\widehat{\Phi}_K(\vect{k}),
		\]
		and inverse FFT to obtain $\vect{g}_K$; add $\vect{g}_b=-\nabla\Phi_b$ and sample $v(R)$ in the mid--plane. Complexity $O(N\log N)$. I set the $k=0$ component of the potential to zero to remove any box--constant offset (this does not affect $\nabla\Phi$).
		
		\paragraph{Axisymmetric speed-up (consistency check).}
		All reported metrics and figures use the 3D--FFT pipeline; the axisymmetric route is a consistency check only.
		
		\subsection{Robustness checks}
		(i) Mock injection/recovery with SPARC-like noise; (ii) distance/inclination perturbations; (iii) bar/warp masks. Code is version-pinned and containerized.
		
		\section{Data and Fitting Protocol}\label{sec:methods}
		\paragraph{Dataset and cuts.}
		Preregistered SPARC subset: inclination $\ge 30^\circ$, quality flag $\le 2$, reliable distances/photometry. Stellar $M/L$ priors are SPS--motivated; see \citep{meidt2014mlr36,schombert2014cmlr}.
		$\Upsilon_\star^{[3.6]}(\mathrm{disk})=0.5\pm 0.1$,
		$\Upsilon_\star^{[3.6]}(\mathrm{bulge})=0.7\pm 0.1$
		These $3.6\,\mu$m priors are standard because this band minimizes dust and traces old stellar populations with relatively small population variance; the adopted dispersions follow \citet{meidt2014mlr36,schombert2014cmlr}. I also report a sensitivity test with broadened priors; widening $\Upsilon_\star$ primarily trades against $\mu$, which our gas--dominated dwarf fold mitigates by construction.
		
		(Gaussian, truncated at $\pm 2\sigma$). Gas from H\,I with factor 1.33.
		Stellar $M/L$ priors are SPS–motivated \citep{meidt2014mlr36,schombert2014cmlr}:
		$\Upsilon_{[3.6],\mathrm{disk}}=0.5\pm0.1$ and
		$\Upsilon_{[3.6],\mathrm{bulge}}=0.7\pm0.1$
		(Gaussian, truncated at $\pm2\sigma$). Gas masses include a factor 1.33 for He.
		These 3.6\,$\mu$m priors minimize dust sensitivity and track old populations with low variance;
		widening them mainly trades against $\mu$, which our gas–dominated dwarf fold mitigates by design.
		\paragraph{Baryonic mass modeling (3D).}
		I deproject SPARC surface maps into a 3D $\rho_b$ as follows.
		\emph{Stellar disk:} exponential in $R$ with catalog scale length $R_d$ and a vertical profile
		${\rm sech}^2(z/h_\star)$ (default $h_\star=0.1\,R_d$ when unavailable), normalized to the
		$3.6\,\mu$m luminosity and $\Upsilon_{[3.6],\mathrm{disk}}$.
		\emph{Bulge:} spherical Sérsic with catalog $(n,R_e)$ or a de Vaucouleurs fallback, with
		$\Upsilon_{[3.6],\mathrm{bulge}}$.
		\emph{Gas:} H\,I surface density with a fixed He factor 1.33; vertical profile ${\rm sech}^2(z/h_{\rm gas})$
		with $h_{\rm gas}=0.3$\,kpc by default and an optional linear flare $h_{\rm gas}(R)=h_{\rm gas}(0)(1+\alpha R/R_d)$
		(\emph{sensitivity only}, not tuned).
		I report a robustness sweep over $(h_\star,h_{\rm gas},\alpha)$ in the supplement; the preregistered fit and gates do not tune them.
		
		\paragraph{Global two-parameter fit.}
		I fit $(\Lmem,\gain)$ by minimizing the velocity-space $\chi^2$ (equivalently RMSE under Gaussian errors) across folds. I report the \emph{median absolute fractional error} (MAFE) and RMSE but do not jointly optimize them; I define $\mathrm{MAFE}=\operatorname{median}_i\!\left(\left|v_{\rm pred}(R_i)-v_{\rm obs}(R_i)\right|/v_{\rm obs}(R_i)\right)$. I report Pearson $r=\mathrm{corr}(\Lmem,\gain)$ and show the \emph{joint posterior / CV scatter} to make any degeneracy explicit.
		
		\paragraph{Baseline for comparison.}
		The comparison baseline uses a per--galaxy \textbf{Burkert} halo (two parameters) under the \emph{same} baryonic priors and CV protocol; an \textbf{NFW} baseline is reported as a sensitivity check.
		
		\paragraph{Degeneracy guardrails.}
		(i) A dedicated CV fold contains gas-dominated dwarfs defined by $M_{\rm gas}/M_\star\ge 1$ (i.e., $f_{\rm gas} \ge 0.5$) and is always present in the 5-fold split; (ii) galaxy-level bootstrap ($B=1000$) for CIs (no point-level leakage).
		
		\paragraph{BTFR.}
		The flat speed $V_f$ is the median of the last $N=4$ valid RC points with $\left|\mathrm{d}\ln v/\mathrm{d}\ln R\right|<0.05$; \emph{the same rule is applied to the halo baseline}. Fit $\log M_b=\alpha\log V_f+\beta$ with intrinsic scatter $\sigma_{\rm int}$.
		
		\paragraph{RAR (parity with literature).}
		RAR is computed as observed total acceleration versus Newtonian baryonic acceleration, \emph{matching the standard literature definition}. I report total scatter and test residual trend vs.\ central surface brightness via a linear slope $t$-test (\textbf{two-sided, $\alpha=0.01$}).
		
		\paragraph{Lensing prediction (no retuning).}
		With $(\Lmem,\gain)$ fixed by RCs, compute $\Delta\Sigma(R)$ and $\gamma_t(R)$ for stacks in \textbf{DES Y3}, \textbf{HSC PDR2}, and \textbf{KiDS-1000} redshift/mass bins (catalog/binning preregistered). The baseline for the variance-reduction gate is baryons--only potential ($\mu=0$) + per-galaxy NFW with a standard mass–concentration prior \citep{dutton2014mcr}, fit per galaxy under the same baryonic priors.
		
\paragraph{Cosmology for $\Sigma_{\rm crit}$.}
Adopt a flat Planck–like cosmology \citep{planck2018vi} (e.g., $\Omega_m=0.315$, $\Omega_\Lambda=0.685$, $h=0.674$) when computing $\Sigma_{\rm crit}$ for all stacks.

		\section{Results}\label{sec:results}
		\subsection{Rotation curves}
		Show a 10-galaxy grid; report CV metrics and Pearson $r$ for $\mathrm{corr}(\Lmem,\gain)$ with the joint scatter plot.
		
		\subsection{BTFR and RAR}
		Report $\alpha$, $\sigma_{\rm int}$ (BTFR) and RAR scatter; confirm no significant trend vs.\ surface brightness.
		
		\subsection{Lensing stacks}
		Compare predicted vs.\ observed $\Delta\Sigma(R)$ or $\gamma_t(R)$; report the preregistered residual-variance reduction relative to the baseline and its significance. I will also stratify lensing stacks by posterior quartiles of $L$; the model predicts a monotonic increase of $\Delta\Sigma(R)$ at $R\sim 30$--$200$\,\si{\kpc} with $L$.
		\paragraph{Planned discriminants (non-gating).}
		I will compare the predicted vertical force law $K_z(R,z)$ against available stellar
		$\sigma_z$ measurements, and test for external-field effects by contrasting otherwise similar
		disks in low/high host-field environments. H1 predicts no intrinsic EFE beyond that encoded in $\rho_b$.
		
		\section{Falsification Gates (preregistered)}\label{sec:falsification}
		\begin{center}
			\begin{tabular}{@{}lcc@{}}
				\toprule
				\textbf{Gate} & \textbf{Pass} & \textbf{Fail}\\
				\midrule
				RC residuals & MAFE $\le 10\%$ and CV-RMSE $\le$ baseline\,$+5\%$ & otherwise\\
				BTFR & $\alpha\in[3.8,4.2]$ and $\sigma_{\rm int}\le 0.12$ dex & otherwise\\
				RAR & total scatter $\le 0.13$ dex and no trend ($\alpha\!=\!0.01$) & otherwise\\
				Lensing (prediction)\footnotemark & $\ge 10\%$ residual-variance reduction at $2\sigma$ vs.\ baseline & otherwise\\
				\bottomrule
			\end{tabular}
			\paragraph{Model comparison rationale.}
			The CV-RMSE margin of $+5\%$ relative to the per-galaxy halo baseline is our parsimony buffer:
			a two-parameter sample-global model that is within $5\%$ in predictive RMSE is preferred on
			complexity grounds. As a sensitivity, I also report sample-aggregated $\Delta{\rm BIC}$
			between H1 and the baseline; $\Delta{\rm BIC}\lesssim -10$ is treated as strong evidence for H1.
			
		\end{center}
		\footnotetext{Relative to the baseline defined in \cref{sec:methods}: baryons--only ($\mu=0$) plus per--galaxy NFW with a standard mass–concentration prior.}
		\section{Physical Motivation and Theoretical Context}\label{sec:motivation}
		\paragraph{Why a convolution?}
		Nonlocal response appears across physics when cumulative structure leaves long--range imprints (screening, memory, effective media). I model such imprinting phenomenologically as a \emph{cored, finite} convolution of the luminous mass, building one field for both stars and photons.
		
		\paragraph{Kernel choice and parsimony.}
		I adopt two cored, positive kernels with finite effective mass: Plummer (\cref{eq:Kpl}) and an exponentially softened core (\cref{eq:KexpCore}). Cored profiles avoid hard--coding flat $1/r$ tails (protecting BTFR scaling) while allowing quasi--flat segments over finite radii. I preregister these two as the primary test and report a \emph{robustness check} across a one--parameter cored family in the supplement (not used for gating).
		
		\paragraph{Interpretation of $(L,\mu)$.}
		$L$ [kpc] is a global correlation length that spreads baryonic influence beyond the bright thin disk; $\mu$ [--] is a dimensionless global gain. H1 does not claim these are fundamental constants; if the gates are passed, the measured $(L,\mu)$ become targets for deeper theory.
		
		\paragraph{Sign of $\rho_{\rm eff}$ and compensation.}
		Because $\rho_{\rm eff}\!\propto\!\nabla^2\Phi_{\rm eff}$ of a smoothed potential, local sign changes (compensation) can occur even with positive $K$. I therefore (i) require physically sensible \emph{projected} lensing profiles and (ii) report a nonnegativity diagnostic (App.~\ref{app:lensing}). This keeps interpretability honest while letting the data decide.
		
		\paragraph{Context with prior ideas.}
		Our approach is phenomenological and distinct from (i) nonlocal gravity kernels fit per galaxy \citep{rahvar2014_nonlocal_rotation} and (ii) luminous--convolution fits with per--galaxy freedom \citep{cisneros2014_lcm}. It also differs from MOND/AQUAL and TeVeS (Sec.~\ref{sec:related_mond}): H1 introduces a \emph{length} scale $L$ via a baryon--built convolution rather than a universal acceleration $a_0$ or a new dynamical field for inertia.
		
		\section{Related Work and Distinctions}\label{sec:related}
		I differ from (i) nonlocal-gravity kernels with tunable functions \citep{rahvar2014_nonlocal_rotation} and (ii) luminous-convolution phenomenology fitted per galaxy \citep{cisneros2014_lcm}: I use a \emph{cored, finite} kernel, fit \emph{one} $(\Lmem,\gain)$ globally, and test RCs, BTFR \citep{lelli2016sparc}, RAR \citep{mcgaugh2016rar}, plus \emph{out-of-sample} lensing with preregistered gates.
		While MOND posits a universal acceleration scale $a_0$, H1 posits a universal length scale $L$; one potential $\Phi_{\rm eff}$ then governs both stars (via $-\nabla\Phi_{\rm eff}$) and photons (via the Poisson/projected–Hessian route), enabling a direct out–of–sample lensing prediction.
		
		\paragraph{Inner--slope diversity (secondary).}
		I compare the distribution of inner logarithmic slopes ${\rm d}\ln v/{\rm d}\ln R$ over $0.5$--$1.5$ disk scale lengths to SPARC values and report a Kolmogorov--Smirnov distance (non--gating).
		
\paragraph{Relation to MOND/TeVeS and other alternatives.}
While both our framework and MOND reproduce the tight RAR, the control scales and constructions differ.
MOND introduces a universal \emph{acceleration} $a_0$ and modifies the low-acceleration force law; relativistic completions (e.g., TeVeS) add fields to mediate lensing.
By contrast, H1 postulates a universal \emph{length} $L$ and a dimensionless gain $\mu$ to build a \emph{baryon–convolved} $\Phi_{\rm eff}$ with Newtonian $1/r^2$ asymptotics.
Once $\Phi_{\rm eff}$ is formed from the luminous map, galaxy–galaxy lensing follows via the standard Poisson/GR route without per–galaxy halos or extra fields.
H1 does not predict a canonical external–field effect; any environment dependence enters only through $\rho_b$ and fixed sample cuts.
I also differ from nonlocal-gravity kernels fit per galaxy and luminous-convolution phenomenology by fitting a single $(L,\mu)$ globally and preregistering RC/BTFR/RAR and out-of-sample lensing gates \citep{rahvar2014_nonlocal_rotation,cisneros2014_lcm,lelli2016sparc,mcgaugh2016rar}.
For broader context see \citep{famaey2012mondreview} and \citet{bekenstein2004teves}.

		
		\section{Discussion and Limitations}\label{sec:discussion}
		Bars/warps and distance systematics can bias RCs; dwarfs with extreme gas fractions test cored-kernel assumptions. Allowing per-galaxy freedom would improve fits but harm identifiability, intentionally avoided. Follow-up H1b will add a relaxation scale $\tau$ to test look-back lensing and merger activity.
		\paragraph{Equivalence principles.}
		Within H1, the same $\Phi_{\rm eff}$ acts on all tracers, so the Weak Equivalence Principle is respected in the phenomenology tested here. I do not address the Strong Equivalence Principle or cosmological backreaction; those require a relativistic completion and are beyond H1's scope.
		
		\paragraph{Scope: clusters and ellipticals.}
		H1 targets rotationally supported disks. Pressure--supported ellipticals and galaxy clusters, where kinematics and lensing probe different regimes, are reserved for follow--up tests with the same $(L,\mu)$ to assess universality.
		
		\paragraph{Reading the best--fit $(L,\mu)$.}
		If the gates are passed, the inferred global $L$ may encode a typical correlation length of baryon--gravity coupling in disks, and $\mu$ a dimensionless gain. I will test correlations between $L$ and outflow/vertical structure proxies (preplanned, non--tuning) to probe physical origin.
		
		
		\section{Conclusion}\label{sec:conclusion}
		A two-parameter, baryon-built effective potential provides a stringent, falsifiable account of disk-galaxy kinematics. A positive lensing prediction without retuning would argue that much of the discrepancy can be encoded in a sample-global, cored convolution with the luminous matter.
		
		\paragraph{Code and Data Availability.}
		I will release a Dockerized reproducer (fixed CV folds, priors, FFT kernels, and pass-fail gates) upon submission.
		
		% =========================================================
		% APPENDICES
		% =========================================================
		\appendix
		
		\section{Units and basic behavior}\label{app:units}
		\subsection*{A.1 Units (single source of truth)}
		
		\subsection*{A.2 Qualitative tails}
		For compact mass, $K_{\text{Plummer}}\sim (4\pi r^2)^{-1}$ at large $r$, so the added field outside the light declines as $1/r^2$. Over observed finite ranges, the superposition with $\Phi_b$ can produce quasi-flat segments; whether it does so across a sample is addressed empirically by our gates. (Exactly flat asymptotes would follow from $K\!\propto\!1/r$, not adopted here.)
		
		\section{Activation window for vertical coupling (sensitivity only)}\label{app:thin}
		To test sensitivity to thin--disk structure while preserving the isotropy of $U$, I keep the kernel unchanged and apply a fixed, normalized vertical window to the \emph{baryonic density} when forming $\Phi_{\rm eff}$:
		\[
		\rho_b^{\rm eff}(R,z) \;=\; \rho_b(R,z)\,W_z(z;h),
		\qquad
		W_z(z;h)=\frac{1}{\sqrt{2\pi}\,h}\,e^{-z^2/2h^2},
		\]
		with \emph{fixed} $h=\SI{0.3}{\kpc}$ (no new fit DOF). $W_z$ integrates to 1, so mid--plane units and the convolution definition of $\Phi_{\rm eff}$ are preserved. I recompute RC residuals with $\rho_b^{\rm eff}$ for sensitivity only; the preregistered two--parameter fit and gates are unchanged.
		\paragraph{Interpretation of $\rho_{\rm eff}$.}
		$\rho_{\rm eff}$ is an \emph{effective} source defined by inverting Poisson’s equation for the modeled $\Phi_{\rm eff}$; local sign changes can occur and should not be interpreted as literal negative mass. Lensing is computed from $\Phi_{\rm eff}$ itself, and $\rho_{\rm eff}$ is used only as a diagnostic decomposition.
		
		\section{Lensing from the same potential}\label{app:lensing}
I map the same $\Phi_{\rm eff}$ to lensing under the Newtonian Poisson closure,
$\nabla^2\Phi_{\rm eff}=4\pi G(\rho_b+\rho_{\rm eff})$,
treating $\rho_{\rm eff}$ as an \emph{effective} source field to be tested observationally.
		
		\textbf{(ii) Surface density and shear.} Project along the line of sight to get
		\[
		\Sigma_{\rm eff}(R)=\int \rho_{\rm eff}(R,l)\,\dd l,\quad
		\Delta\Sigma(R)=\bar{\Sigma}(<R)-\Sigma_{\rm eff}(R),\quad
		\gamma_t(R)=\frac{\Delta\Sigma(R)}{\Sigma_{\rm crit}}.
		\]
		Equivalently, one may form the 2D lensing potential from the line-of-sight integral of $\phieff$ and take its projected second derivatives; both routes are consistent and implemented numerically from the same $\phieff$ grid.
\paragraph{Nonnegativity diagnostics (reported, non--gating).}
I report (i) the fraction of projected radii with $\Sigma_{\rm eff}(R)<0$ and (ii) the 3D compensation ratio
\[
f_{\rm neg}^{\rm 3D} \equiv \frac{\int_{\mathcal V}\!\max\{0,-\rho_{\rm eff}\}\,{\rm d}^3\!r}
{\int_{\mathcal V}\!\max\{0,\rho_{\rm eff}\}\,{\rm d}^3\!r},
\]
over $\mathcal V$ spanning the analysis box. Physically acceptable models should have $\Sigma_{\rm eff}(R)\ge 0$ within uncertainties and small $f_{\rm neg}^{\rm 3D}$ (I flag $f_{\rm neg}^{\rm 3D}>0.2$). These are diagnostics only and do not constitute preregistered gates.
\noindent\textit{Modeling postulate.} In H1, the same scalar $\Phi_{\rm eff}$ that governs kinematics is assumed to source weak lensing via the Newtonian Poisson mapping; this is an empirical hypothesis adjudicated by the preregistered gates.

		\section{Reproducibility checklist}\label{app:reprod}
		\begin{itemize}
			\item SPARC subset list; CV folds (seed=42); galaxy-level bootstraps ($B=1000$).
			\item $M/L$ priors and distance/inclination uncertainty models.
			\item FFT grid sizes; padding factor 2; kernel truncation $R_{\rm trunc}=6\,\Lmem$; $R_{\max}\ge 1.5\,R_{\rm trunc}$.
			\item Mock injection/recovery scripts and evaluation metrics.
			\item Pass/fail thresholds mirrored from \cref{sec:falsification}.
		\end{itemize}
		\paragraph{Code and Data Availability.}
  All code and the LaTeX source are archived at Zenodo, DOI: \href{https://doi.org/10.5281/zenodo.16967259}{10.5281/zenodo.16967259}.

		% ----------------- Bibliography -----------------
		\bibliographystyle{unsrtnat}
		\bibliography{references}
		
	\end{document}
